---
title: "An introduction to the cassowaryr package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vaast}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# About 

<!-- badges: start -->
<!-- badges: end -->

The `cassowaryr` package provides functions to compute scagnostics on pairs of numeric variables in a data set.

The term __scagnostics__ refers to scatter plot diagnostics, originally described by John and Paul Tukey. This is a collection of techniques for automatically extracting interesting visual features from pairs of variables. This package is an implementation of graph theoretic scagnostics developed by Wilkinson, Anand, and Grossman (2005) in pure R.  

# Installation

The package can be installed from CRAN using 

> ```install.packages("cassowaryr")```

and from GitHub using 

> ```remotes::install_github("numbats/cassowaryr")```

to install the development version.

# Examples 

```{r setup}
library(cassowaryr)
library(ggplot2)
library(dplyr)

# pick examples
exampledata <- datasaurus_dozen %>%
  filter(dataset==c("v_lines", "circle", "dots", "away")) 

#plot them
exampledata %>%
  ggplot(aes(x=x,y=y, colour=dataset))+
  geom_point() +
  facet_wrap("dataset") +
  theme_minimal() +
  theme(legend.position = "none")
```
We can calculate a select group of scagnostic for every group in a dataset.
```{r scag}
#compute scagnostics
scag_data <- exampledata %>%
  group_by(dataset) %>%
  summarise(calc_scags(x,y, scags=c("clumpy2", "outlying", "monotonic"))) 

scag_data %>%
  knitr::kable(digits=4, align="c")
```

We can also a select group of scagnostics for every pairwise variable combination of a data set.

```{r}
data(pk)
calc_scags_wide(pk[,2:5], scags=c("outlying","clumpy","monotonic")) %>%
  knitr::kable(digits=4, align="c")
```

We can use either of these generated data sets to find the top scagnostic for each variable, or the top variable for each scagnostic. This is an example of the top scagnsotic for each pair.

```{r}
scag_data %>%
  top_pairs() %>%
  knitr::kable(digits=4, align="c")
```
